2.2: Random Variables and Distributions

### 2019-09-19 static/slides/08-slides.Rmd
Fix slides 8

@@

$$\begin{align*}\sigma^2_X &= E[(X-\mu_X)^2]\\
&=\sum_{i=1}^n(x_i-\mu_X)^2p_i\\ \end{align*}$$

@@

@@

.center[
$P(0 \leq X \leq 2)=$ area under the curve between 0 and 2
]
]

@@

@@

.center[
$P(X \leq 2)=$ area under the curve to the left of 2
]
]

@@

@@

```{r, fig.retina=3}
ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  labs(x = "X",
       y = "Probability")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68\%

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68\%

- $P(\mu-2\sigma \leq X \leq \mu+2\sigma) \approx$ 95\%

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68\%

- $P(\mu-2\sigma \leq X \leq \mu+2\sigma) \approx$ 95\%

- $P(\mu-3\sigma \leq X \leq \mu+3\sigma) \approx$ 99.7\%
]

@@

- The .onfire[standard] normal distribution has mean 0 and standard deviation 1

@@

# Standardizing Variables: Example I

@@

```{r, fig.retina=3, fig.align="center"}
shade_neg4_to_1 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < -4 | x >1] <- NA
    return(y)
}

@@

```{r, fig.retina=3, fig.align="center"}
shade_1_to_4 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < 1 | x >4] <- NA
    return(y)
}

@@

```{r, fig.retina=3, fig.align="center"}
shade_neg1_to_1 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < -1 | x >1] <- NA
    return(y)
}

@@

```{r, echo=T}
# subtract two left tails!
pnorm(85, # larger number first!
      mean = 75,
      sd = 10,
      lower.tail = TRUE)- # looking to left
  pnorm(65, # smaller number second!
        mean = 75,
        sd = 10,
        lower.tail = TRUE) #looking to left
```
]

>>>>

class: inverse, center, middle
# Discrete Random Variables

@@

$$\begin{align*}\sigma^2_X &= E[(x_i-\mu_X)^2]\\
&=\sum_{i=1}^n(x_i-\mu_X)^2p_i\\ \end{align*}$$

@@

.content-box-green[
.green[**Example**]: P(0 \leq X \leq 2)
]

@@

@@

.content-box-green[
.green[**Example**]: P(X \leq 2)
]
]

@@

@@

.content-box-green[
.green[**Example**]: P(X \geq 2) = 1 - P(X \leq 2)
]
]

@@

```{r, fig.retina=3}
ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  scale_x_continuous(breaks=seq(-4,4,1))+
  labs(x = "X",
       y = "Probability")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68%

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68%

- $P(\mu-2\sigma \leq X \leq \mu+2\sigma) \approx$ 95%

@@

- $P(\mu-1\sigma \leq X \leq \mu+1\sigma) \approx$ 68%

- $P(\mu-2\sigma \leq X \leq \mu+2\sigma) \approx$ 95%

- $P(\mu-3\sigma \leq X \leq \mu+3\sigma) \approx$ 99.7%

- **68/95/99.7%** of observations fall within **1/2/3 standard deviations** of the mean
]

@@

- The .onfire[standard] normal distribution (often referred to as **Z**) has mean 0 and standard deviation 1

@@

# Standardizing Variables:

- We can take any normal distribution (for any $\mu, \sigma)$ and **standardize** it to the standard normal distribution by taking the .shout[Z-score] of any value, $x_i$:

$$Z=\frac{x_i-\mu}{\sigma}$$

- Subtract any value by the distribution's mean and divide by standard deviation

- $Z$: number of standard deviations $x_i$ value is away from the mean
]

.pull-right[
```{r, fig.retina=3}
normal_pdf
```
]

# Standardizing Variables: Example II

@@

```{r, fig.retina=3, fig.height=5, fig.align="center"}
shade_neg4_to_1 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < -4 | x >1] <- NA
    return(y)
}

@@

```{r, fig.retina=3, fig.height=5, fig.align="center"}
shade_1_to_4 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < 1 | x >4] <- NA
    return(y)
}

@@

```{r, fig.retina=3, fig.height=5, fig.align="center"}
shade_neg1_to_1 <- function(x) {
    y <- dnorm(x, mean = 0, sd = 1)
    y[x < -1 | x >1] <- NA
    return(y)
}

@@

```{r, echo=T}
# subtract two left tails!
pnorm(85, # larger number first!
      mean = 75,
      sd = 10,
      lower.tail = TRUE) - # looking to left, & SUBTRACT
  pnorm(65, # smaller number second!
        mean = 75,
        sd = 10,
        lower.tail = TRUE) #looking to left
```
]
